Music adaptation: Analysis

### 2022-10-14 analysis/baseline.Rmd
update analysis with baseline
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@

```{r}
aov_ez(
  data = ioi,
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

@@

```{r}
aov_ez(
  data = ioi[Category == "both" | Category == "none" | Category == "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

@@

```{r}
aov_ez(
  data = ioi,
  id = "SubNr",
  dv = "SD",
  within = c("Category")
)
```

@@

```{r kot-all-plot, echo = FALSE}
ggplot(kot[Subcomponent1 == "Legato"], aes(x = Articulation, y = Mean, color = Dynamics))+ 
  geom_boxplot(position = position_dodge2(preserve = "single")) +
  #geom_point(position = position_dodge(0.75)) +
  facet_grid( ~ Subcomponent1) +
  theme_pubr()

ggplot(kot[Subcomponent1 == "Staccato"], aes(x = Articulation, y = Mean, color = Dynamics))+ 
  geom_boxplot(position = position_dodge2(preserve = "single")) +
  #geom_point(position = position_dodge(0.75)) +
  facet_grid( ~ Subcomponent1) +
  theme_pubr()

@@

```{r}
aov_ez(
  data = kot[Subcomponent1 == "Staccato" & Category != "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Articulation", "Dynamics")
)
```
2) Category: All

```{r}
sta_ca_all_aov <- aov_ez(
  data = kot[Subcomponent1 == "Staccato"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
sta_ca_all_aov

pairs(emmeans(sta_ca_all_aov, "Category"), adjust = "tukey")
```

@@

```{r}
sta_ca_par_aov <- aov_ez(
  data = kot[Subcomponent1 == "Staccato" & Category == "both" | Subcomponent1 == "Staccato" & Category == "dyn_only" | Subcomponent1 == "Staccato" & Category == "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
sta_ca_par_aov

pairs(emmeans(sta_ca_par_aov, "Category"), adjust = "tukey")
```
# Dynamics
## 1. Individual

@@

```{r vel-all-plot,  echo = FALSE}
ggplot(vel, aes(x = Dynamics, y = Mean, color = Articulation))+ 
  geom_boxplot(position = position_dodge2(preserve = "single")) +
  geom_point(binaxis = "y", stackdir = "center", position = position_dodge(0.75)) +
  facet_grid( ~ Subcomponent2) +
  theme_pubr()

@@

ggline(vel_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Note Nr", ylab = "Velocity (0-127)", title = "Velocity") + scale_x_continuous(breaks = seq(1,72,1))
```

1) Forte

```{r}
aov_ez(
  data = vel[Subcomponent2 == "Forte"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

2) Piano

```{r}
aov_ez(
  data = vel[Subcomponent2 == "Piano"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

@@

## 2. All - group
```{r vel-diff-all, echo = FALSE}
# group mean
vel_diff_all <- vel_diff[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Category, Subcomponent2)]
vel_diff_all
```

@@

```{r vel-diff-all-plot,  echo = FALSE}
ggplot(vel_diff, aes(x = Dynamics, y = Mean, color = Articulation))+ 
  geom_boxplot(position = position_dodge2(preserve = "single")) +
  geom_point(binaxis = "y", stackdir = "center", position = position_dodge(0.75)) +
  facet_grid( ~ Subcomponent2) +
  theme_pubr()

@@

1) FtoP

```{r}
aov_ez(
  data = vel_diff[Subcomponent2 == "FtoP"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

2) PtoF

```{r}
aov_ez(
  data = vel_diff[Subcomponent2 == "PtoF"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

>>>>

## Experimental Design

- Independent variables: Articulation (2 levels: Present or Absent) x Dynamics (2 levels: Present or Absent)
- Dependent variables: Interonset intervals (IOIs), Key-overlap time (KOT), Key velocity (KV) and KV Difference

1. Main predictions ([preregistration](https://osf.io/qbzwt))

- If students performed the piece without articulation (i.e., only dynamics is implemented in the recordings), participants should exaggerate articulation more (i.e., producing longer legato and shorter staccato).
- If students performed the piece without dynamics (i.e., only articulation is implemented in the recordings), participants should exaggerate dynamics more (i.e., producing louder forte and softer piano).
- If students performed the piece without articulation and dynamics (i.e., neither articulation nor dynamics is implemented in the recordings), participants should exaggerate both articulation and dynamics more.

2. Excluded participants

- NA

3. Stats packages

- [afex](https://github.com/singmann/afex) for ANOVAs

@@

```{r}
ioi_ca_all_aov <- aov_ez(
  data = ioi,
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)

ioi_ca_all_aov$anova_table
pairs(emmeans(ioi_ca_all_aov, "Category"), adjust = "tukey")
```

@@

```{r}
ioi_ca_par_aov <- aov_ez(
  data = ioi[Category == "both" | Category == "none" | Category == "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)

ioi_ca_par_aov$anova_table
pairs(emmeans(ioi_ca_par_aov, "Category"), adjust = "tukey")
```

@@

```{r}
var_stats <- aov_ez(
  data = ioi,
  id = "SubNr",
  dv = "SD",
  within = c("Category")
)

var_stats$anova_table
```

@@

```{r kot-all-plot, echo = FALSE}
ggplot(kot[Subcomponent1 == "Legato"], aes(x = Articulation, y = Mean, color = Dynamics)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
  labs(y = "KOT (ms)", title = "KOT - Legato") +
  theme_pubr()

ggplot(kot[Subcomponent1 == "Staccato"], aes(x = Articulation, y = Mean, color = Dynamics)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
  labs(y = "KOT (ms)", title = "KOT - Staccato") +
  theme_pubr()

@@

```{r}
sta_all_aov <- aov_ez(
  data = kot[Subcomponent1 == "Staccato" & Category != "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Articulation", "Dynamics")
)

sta_all_aov$anova_table
pairs(emmeans(sta_all_aov, ~Dynamics|Articulation), adjust = "tukey")
```

```{r}
sta_ca_all_aov <- aov_ez(
  data = kot[Subcomponent1 == "Staccato"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)

sta_ca_all_aov$anova_table
pairs(emmeans(sta_ca_all_aov, "Category"), adjust = "tukey")
```

@@

```{r}
sta_ca_par_aov <- aov_ez(
  data = kot[Subcomponent1 == "Staccato" & Category == "both" | Subcomponent1 == "Staccato" & Category == "dyn_only" | Subcomponent1 == "Staccato" & Category == "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)

sta_ca_par_aov$anova_table
pairs(emmeans(sta_ca_par_aov, "Category"), adjust = "tukey")
```

# Dynamics
## 1. Individual

@@

```{r vel-all-plot,  echo = FALSE}
ggplot(vel[Subcomponent2 == "Forte"], aes(x = Dynamics, y = Mean, color = Articulation))+ 
    geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
  labs(y = "Key Velocity (0-127)", title = "KV - Forte") +
  theme_pubr()

ggplot(vel[Subcomponent2 == "Piano"], aes(x = Dynamics, y = Mean, color = Articulation))+ 
    geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
  labs(y = "Key Velocity (0-127)", title = "KV - Piano") +
  theme_pubr()

@@

ggline(vel_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Note Nr", ylab = "Key Velocity (0-127)", title = "KV") + scale_x_continuous(breaks = seq(1,72,1))
```

## 4. Stats
### Forte

```{r}
for_all_aov <- aov_ez(
  data = vel[Subcomponent2 == "Forte" & Category != "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Articulation", "Dynamics")
)

for_all_aov$anova_table
pairs(emmeans(for_all_aov, ~Articulation|Dynamics), adjust = "tukey")
```

```{r}
for_ca_all_aov <- aov_ez(
  data = vel[Subcomponent2 == "Forte"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)

for_ca_all_aov$anova_table
pairs(emmeans(for_ca_all_aov, "Category"), adjust = "tukey")
```

3) Category: both vs. art_only vs. baseline

```{r}
for_ca_par_aov <- aov_ez(
  data = vel[Subcomponent2 == "Forte" & Category == "both" | Subcomponent2 == "Forte" & Category == "art_only" | Subcomponent2 == "Forte" & Category == "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)

for_ca_par_aov$anova_table
pairs(emmeans(for_ca_par_aov, "Category"), adjust = "tukey")
```

### Piano

```{r}
pia_all_aov <- aov_ez(
  data = vel[Subcomponent2 == "Piano" & Category != "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Articulation", "Dynamics")
)

pia_all_aov$anova_table
pairs(emmeans(pia_all_aov, ~Articulation|Dynamics), adjust = "tukey")
```

```{r}
pia_ca_all_aov <- aov_ez(
  data = vel[Subcomponent2 == "Piano"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)

pia_ca_all_aov$anova_table
pairs(emmeans(pia_ca_all_aov, "Category"), adjust = "tukey")
```

3) Category: both vs. art_only vs. baseline

```{r}
pia_ca_par_aov <- aov_ez(
  data = vel[Subcomponent2 == "Piano" & Category == "both" | Subcomponent2 == "Piano" & Category == "art_only" | Subcomponent2 == "Piano" & Category == "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)

pia_ca_par_aov$anova_table
pairs(emmeans(pia_ca_par_aov, "Category"), adjust = "tukey")
```

@@

## 2. Group

```{r vel-diff-all, echo = FALSE}
# group mean
vel_diff_all <- vel_diff[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Category, Subcomponent2)]
vel_diff_all
```

@@

```{r vel-diff-all-plot,  echo = FALSE}
ggplot(vel_diff[Subcomponent2 == "FtoP"], aes(x = Dynamics, y = Mean, color = Articulation))+ 
  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
  labs(y = "KV Difference (-127-127)", title = "KV Difference - Forte to Piano") +
  theme_pubr()

ggplot(vel_diff[Subcomponent2 == "PtoF"], aes(x = Dynamics, y = Mean, color = Articulation))+ 
  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
  labs(y = "KV Difference (-127-127)", title = "KV Difference - Piano to Forte") +
  theme_pubr()

@@

## 4. Stats
### FtoP

```{r}
ftop_all_aov <- aov_ez(
  data = vel_diff[Subcomponent2 == "FtoP" & Category != "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Articulation", "Dynamics")
)

ftop_all_aov$anova_table
pairs(emmeans(ftop_all_aov, ~Articulation|Dynamics), adjust = "tukey")
```

```{r}
ftop_ca_all_aov <- aov_ez(
  data = vel_diff[Subcomponent2 == "FtoP"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)

ftop_ca_all_aov$anova_table
pairs(emmeans(ftop_ca_all_aov, "Category"), adjust = "tukey")
```

3) Category: both vs. art_only vs. baseline

```{r}
ftop_ca_par_aov <- aov_ez(
  data = vel_diff[Subcomponent2 == "FtoP" & Category == "both" | Subcomponent2 == "FtoP" & Category == "art_only" | Subcomponent2 == "FtoP" & Category == "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)

ftop_ca_par_aov$anova_table
pairs(emmeans(ftop_ca_par_aov, "Category"), adjust = "tukey")
```

### PtoF

```{r}
ptof_all_aov <- aov_ez(
  data = vel_diff[Subcomponent2 == "PtoF" & Category != "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Articulation", "Dynamics")
)

ptof_all_aov$anova_table
pairs(emmeans(ptof_all_aov, ~Articulation|Dynamics), adjust = "tukey")
```

```{r}
ptof_ca_all_aov <- aov_ez(
  data = vel_diff[Subcomponent2 == "PtoF"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)

ptof_ca_all_aov$anova_table
pairs(emmeans(ptof_ca_all_aov, "Category"), adjust = "tukey")
```

3) Category: both vs. art_only vs. baseline

```{r}
ptof_ca_par_aov <- aov_ez(
  data = vel_diff[Subcomponent2 == "PtoF" & Category == "both" | Subcomponent2 == "PtoF" & Category == "art_only" | Subcomponent2 == "FtoP" & Category == "baseline"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)

ptof_ca_par_aov
pairs(emmeans(ftop_ca_par_aov, "Category"), adjust = "tukey")
```
